The Etiology of Cybersecurity

### 2022-11-24 reports/cimss_2022/main-cut.tex

Following \autocite{Santaca2016abf}, we define a system as a hierarchy of agents. Furthermore, we model an agent as a meronomy (an
hierarchy of part-whole relations) over the aforementioned constituents 
(assertions, beliefs, and facts), based on a standard definition of mereology
(i.e. based on the definition of parthood relation between \emph{parts}).  Due to
the necessity of considering different relations between parts (as we will show afterwards)
we extend the mereology to a
mereo-topology \autocite{Smith1996mereotopology,Varzi1994mereotopology,Rachavelpula2017mereotopology},
considering the relations in Tab.~\ref{tab:rcc358}.  For the sake of
readability, we use the term \emph{region} both to refer to a mereological part
and to a topological region.  Our aim is to create a meronomy (hierarchy of
part-whole relations) instead of the taxonomies (categorization based on
discrete sets) such as the one provided in \autocite{NIST2020NVD,CVE}
so that we don't need to rely on a scoring system (such as the CVSS) to assign
a quantitative evaluation of the cybersecurity of each entry. Instead, we want
a precise calculation of the number of insecure configurations of a system to
emerge from the mathematical formulation of our cybersecurity hypothesis.
A mereotopology, as defined in \autocite{Rachavelpula2017mereotopology},
is a mathematical structure where the basic relation between regions
is the reflexive and symmetric relation \emph{Connects With},
that we use to order a universe of agents
$\agentuniverse$ (see in Tab.~\ref{tab:rcc358}).  We use the Region Connection
Calculus (RCC), as defined in \autocite{bennettLogics,improvingRCC}, to provide
an axiomatization of the mereo-topological concepts. In its broader definition,
the RCC theory is composed by eight axioms, and is known as RCC8 \autocite{Grutter2008rcc}. Using
RCC5 instead of RCC8 prevents us from considering tangential connections
between spatial regions. However, tangential connections in RCC8 can be
considered as special cases of the more general spatial relations considered in
RCC5.
In Tab.~\ref{tab:rcc}, we summarize the axioms of the RCC (see, e.g., \autocite{Grutter2008rcc}).  We can now define a system
over the mereotopology using the RCC calculus, as follows, where $\rcc(X,Y)$ on
two generic regions $X,Y$ represents one of the possible RCC relations between
$X$ and $Y$. We note that all the RCC relations are symmetric with the exception
of those that have an explicit (related) inverse.

@@

As in \autocite{Santaca2016abf}, it follows that, by defining
a system as a fixed number of regions, there exists
an upper-bound to the number of possible configuration of a system, defined by
the possible relations between the different regions.
The general formula to calculate the number of different types of agents is
$r^{\binom{n}{k}}$, where $r$ is the number of relations with arity $k$,
between $n$ different regions. In our case, $\binom{n}{k}=3$ since we consider $3$ regions 
($\assertionRegion$, $\beliefRegion$, and $\factRegion$), and all the relations
considered in the RCC are binary.  Hence, we have up to 125
different types of agents but only 54 of the 125 (as showed in
\autocite{improvingRCC}) combinations are topologically correct. 
For RCC5 there are $5^3=125$ theoretical combinations but
only 54 are correct with respect to the axioms.

>>>>

Following \autocite{Santaca2016abf}, we define a system as a hierarchy of agents. Furthermore, we model an agent as a meronomy (an
hierarchy of part-whole relations) over the aforementioned constituents 
(assertions, beliefs, and facts), based on a standard definition of mereology
(i.e. based on the definition of parthood relation between \emph{parts}).  Due to
the necessity of considering different relations between parts (as we will show afterwards)
we extend the mereology to a
mereo-topology \autocite{Smith1996mereotopology,Varzi1994mereotopology,Rachavelpula2017mereotopology},
considering the relations in Tab.~\ref{tab:rcc358}.  For the sake of
readability, we use the term \emph{region} both to refer to a mereological part
and to a topological region.  Our aim is to create a meronomy (hierarchy of
part-whole relations) instead of the taxonomies (categorization based on
discrete sets) such as the one provided in \autocite{NIST2020NVD,CVE}
so that we don't need to rely on a scoring system (such as the CVSS) to assign
a quantitative evaluation of the cybersecurity of each entry. Instead, we want
a precise calculation of the number of insecure configurations of a system to
emerge from the mathematical formulation of our cybersecurity hypothesis.
A mereotopology, as defined in \autocite{Rachavelpula2017mereotopology},
is a mathematical structure where the basic relation between regions
is the reflexive and symmetric relation \emph{Connects With},
that we use to order a universe of agents
$\agentuniverse$ (see in Tab.~\ref{tab:rcc358}).  We use the Region Connection
Calculus (RCC), as defined in \autocite{bennettLogics,Grutter2008rcc}, to provide
an axiomatization of the mereo-topological concepts. In its broader definition,
the RCC theory is composed by eight axioms, and is known as RCC8 \autocite{Grutter2008rcc}. Using
RCC5 instead of RCC8 prevents us from considering tangential connections
between spatial regions. However, tangential connections in RCC8 can be
considered as special cases of the more general spatial relations considered in
RCC5.
In Tab.~\ref{tab:rcc}, we summarize the axioms of the RCC (see, e.g., \autocite{Grutter2008rcc}).  We can now define a system
over the mereotopology using the RCC calculus, as follows, where $\rcc(X,Y)$ on
two generic regions $X,Y$ represents one of the possible RCC relations between
$X$ and $Y$. We note that all the RCC relations are symmetric with the exception
of those that have an explicit (related) inverse.

@@

As in \autocite{Santaca2016abf}, it follows that, by defining
a system as a fixed number of regions, there exists
an upper-bound to the number of possible configuration of a system, defined by
the possible relations between the different regions.
The general formula to calculate the number of different types of agents is
$r^{\binom{n}{k}}$, where $r$ is the number of relations with arity $k$,
between $n$ different regions. In our case, $\binom{n}{k}=3$ since we consider $3$ regions 
($\assertionRegion$, $\beliefRegion$, and $\factRegion$), and all the relations
considered in the RCC are binary.  Hence, we have up to 125
different types of agents but only 54 of the 125 (as showed in
\autocite{Grutter2008rcc}) combinations are topologically correct. 
For RCC5 there are $5^3=125$ theoretical combinations but
only 54 are correct with respect to the axioms.
